---
title: "p8105_hw2_yj2580"
author: "yj2580"
date: "9/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
library(dplyr)
```

## Problem 1

```{r}
# Create a dataframe from the Mr. Trash Wheel sheet
trash_df = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1)%>%
#clean names, make variable names reasonable
  janitor::clean_names() %>%
#omit non data entries
  select(-x15, -x16, -x17) %>%
#omit rows that do not include dumpster-specific data
  drop_na(dumpster)  %>%
#round the number of sports balls and transfer them into integers
  mutate(sports_balls = as.integer(round(as.numeric(sports_balls))))


```

```{r}
#Create a dataframe of precipitation data for 2018
pre2018_df = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 3, range = "A2:B14") %>%
  janitor::clean_names() %>%
#add a variable year, convert month to a character variable
  mutate(year = "2018",
         month = month.name)  %>%
#omit rows without precipitation data
  drop_na()

#Create a dataframe of precipitation data for 2017
pre2017_df = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 4, range = "A2:B14") %>%
  janitor::clean_names() %>%
#add a variable year,convert month to a character variable
  mutate(year = "2017",
         month = month.name)  %>%
#omit rows without precipitation data
  drop_na()

#combine precipitation datasets
pre_df = rbind(pre2017_df, pre2018_df)

```

Comments: 
In the Mr. Trash Wheel, we can see the summary of variables from this table: 
```{r, echo = FALSE}
str(trash_df)
```
so there are `r nrow(trash_df)` observations and `r ncol(trash_df)` variables in total.
The key variables for trash data are dumpter number (dumpster), date of collection(year, month, date), amount of total litter (weight_tons, volume_cubic_yards) and litter types (plastic_bottles, polystyrene, cigarette_butts, glass_bottles, grocery_bags, chip_bags, sports_balls, homes_powered).

In the precipitation data in 2017 and 2018, we can see the summary of variables from this table: 
```{r, echo = FALSE}
str(pre_df)
```
so there are `r nrow(pre_df)` observations and `r ncol(pre_df)` variables in total. The key variables includes month, year and the total precipitation amounts.

The total precipitation in 2018 was `r sum(pull(filter(pre_df, year == "2018"),total))`. The median number of sports balls in a dumpster in 2017 was `r median(pull(filter(trash_df, year == "2017"), sports_balls))`.

## Problem 2

```{r}
#create a dataframe
pols_month_df = read_csv("./data/pols-month.csv") %>%
  janitor::clean_names()  %>%
  #seperate date into year, month and day
  separate(mon, c("year", "month", "day")) %>%
  mutate(year = as.integer(year),
         month = as.integer(month),
         day = as.integer(day),
         #replace month number with month name
         month = as.factor(month.name[month]),
         #create a president variable taking values gop and dem
         president = ifelse(prez_gop == "1", "gop", "dem")) %>%
  #remove prez_dem and prez_gop, and the day variable
  select(-day, -prez_gop, -prez_dem)

snp_df = read_csv("./data/snp.csv") %>%
  janitor::clean_names()  %>%
  separate(date, c("month", "day", "year")) %>%
  mutate(year = as.integer(year),
         month = as.integer(month),
         day = as.integer(day),
         month = month.name[month],
         #in order to order by month
         month = factor(month, levels = month.name)) %>% 
  arrange(year, month) %>%
  #organize so that year and month are the leading columns
  select(year, month, everything())

unemploy_df = read_csv("./data/unemployment.csv") %>%
  janitor::clean_names()  %>%
  pivot_longer(
    jan:dec,
    names_to = "month", 
    values_to = "unemployment") %>%
  #change month into full name to keep the same key variable name and the same value
  mutate(month = factor(month, labels = month.name),
         year = as.integer(year))

#merge snp into pols
snp_pols = left_join(pols_month_df, snp_df, by = c("year", "month"))
snp_pols
#merge unemployment into the result
snp_pols_unemploy = left_join(snp_pols, unemploy_df, by = c("year", "month"))
snp_pols_unemploy
```
Comment:
The data “pols-month” contains `r nrow(pols_month_df)` entries with `r ncol(pols_month_df)` columns ranging from 1947 to 2015, related to the number of national politicians who are democratic or republican at any given time. The data "snp" contains `r nrow(snp_df)` entries with `r ncol(snp_df)` columns ranging from 1950 to 2015, related to Standard & Poor’s stock market index. The data "unemployment" contains `r nrow(snp_df)` entries with `r ncol(snp_df)` columns ranging from 1950 to 2015, related to percentage of unemployment in the associated month of the associated year. After merging "snp" into "pols-month" data by month and year, superfluous data on July. 2015 in snp was deleted and the dimension became `r nrow(snp_pols)` entries(the same as "pols-month") with `r ncol(snp_pols)` columns(two new columns from "snp") ranging from 1947 to 2015. After merging "unemployment" into resulting dataframe, superfluous data on some months of 2015 in "unemployment" was deleted and the dimension became `r nrow(snp_pols_unemploy)` entries(still the same as "pols-month") with `r ncol(snp_pols_unemploy)` columns(one new column from "unemployment") ranging from 1947 to 2015. The key variables in this combining dataframe are shown below: 
year, month: date of the count
day: the date of the observation
president: indicator of whether the president was republican or democratic on the associated date (gop = republican, dem = democratic)
gov_gop: the number of republican governors on the associated date
sen_gop: the number of republican senators on the associated date
rep_gop: the number of republican representatives on the associated date
gov_dem: the number of democratic governors on the associated date
sen_dem: the number of democratic senators on the associated date
rep_dem: the number of democratic representatives on the associated date
close: the closing values of the S&P stock index on the associated date
unemployment: percentage of unemployment in the associated month of the associated year

## Problem 3

```{r}
#create popular baby name dataframe
pop_baby_name = read_csv("./data/Popular_Baby_Names.csv") %>%
  janitor::clean_names()  %>%
  #uppercase child's first name to make data look better
  mutate(childs_first_name = toupper(childs_first_name),
         # unify ethnicity name
         ethnicity = ifelse(ethnicity == "ASIAN AND PACI", "ASIAN AND PACIFIC ISLANDER", ethnicity),
         ethnicity = ifelse(ethnicity == "BLACK NON HISP", "BLACK NON HISPANIC", ethnicity),
         ethnicity = ifelse(ethnicity == "WHITE NON HISP", "WHITE NON HISPANIC", ethnicity)) %>%
  #delete duplicate rows
  distinct()

# the rank in popularity of the name “Olivia” as a female baby name over time
filter(pop_baby_name, childs_first_name == "OLIVIA") %>%
  select(year_of_birth, ethnicity, rank) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  ) %>%
  knitr::kable()

# the most popular name among male children over time
filter(pop_baby_name, rank == "1", gender == "MALE") %>%
  select(year_of_birth, ethnicity, childs_first_name ) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  ) %>%
  knitr::kable()

# pull the data of male, white born in 2016
male_whnohis_16 = filter(pop_baby_name, gender == "MALE", ethnicity =="WHITE NON HISPANIC", year_of_birth == "2016")
# produce a scatter plot 
scatterplot = ggplot(male_whnohis_16, aes(y = count, x = rank)) +geom_point() + labs(title = "Number of children with a name against the rank in popularity of that name", x ="rank", y = "number of children")
scatterplot
ggsave("scatterplot.pdf")
```

